---
title: "JFITheme"
output: github_document
---
```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

JFITheme is an R package that provides standardized ggplot2 themes, color scales, and typography for the Jain Family Institute (JFI).

## Installation

To install the `JFITheme` package, use the following commands:

```r
# Install the development version from GitHub
devtools::install_github("JainFamilyInstitute/JFITheme")
```
## Usage
When the package is loaded, custom fonts are automatically registered, so you do not need to manually call `custom_font()`.

```{r}
library(ggplot2)
library(JFITheme)
```

Create a Basic Plot with the JFI Theme
```{r}
# Create a basic plot
jfi_plot(mtcars, aes(x = wt, y = mpg)) +
  geom_point() +
  scale_y_continuous(limits = c(0, 40), expand = c(0, 0)) +
  labs(
    title = "Weight vs MPG",
    subtitle = "Scatter plot of car weight vs. fuel efficiency"
  )
```


Use custom JFI Color Scales and add caption using the `scale_color_jfi()` and `caption_w_logo()` functions
```{r}
# Apply JFI color scales
jfi_plot(mtcars, aes(x = factor(cyl), y = mpg, fill = factor(cyl))) +
  geom_col(width = .75) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  scale_fill_jfi(type = "discrete", palette = "jfi") +
  labs(
    title = "Fuel Efficiency by Cylinder Count",
    subtitle = "A bar plot using JFI discrete color scales",
    x = "Cylinder", y = "Miles per Gallon",
    fill = "Cylinder", 
    caption = caption_w_logo(source = "Data Source",
                             note = "Additional Notes")
  ) + 
  theme(legend.position = "none") # remove legend
```

Example for Sequential Color Scale
```{r}
# Apply JFI continuous color scale
jfi_plot(mtcars, aes(x = wt, y = mpg, color = hp)) +
  geom_point() +
  scale_y_continuous(limits = c(0, 35), expand = c(0, 0)) +
  scale_color_jfi(type = "sequential", palette = "jfi") +
  labs(
    title = "Weight vs MPG by HP with Continuous Color Scale",
    subtitle = "Using JFI sequential color scale",
    color = "Horsepower",
    caption = caption_w_logo(source = "Data Source")
  ) +
  guides(color = guide_colorbar(barwidth = 10, barheight = .7))
```


Example for Divergent Color Scale
```{r}
# Load the airquality dataset and create a temperature deviation variable
# Load the airquality dataset
data(airquality)

airquality$TempDev <- airquality$Temp - mean(airquality$Temp, na.rm = TRUE)

# Apply JFI divergent color scale using temperature deviation
jfi_plot(airquality, aes(x = Wind, y = Temp, color = TempDev)) +
  geom_point(size = 2.5) +
  scale_color_jfi(type = "divergent", palette = "jfi", midpoint = 0) +
  labs(
    title = "Wind vs Temp with Divergent Color Scale",
    subtitle = "Divergence based on temperature deviation from the mean",
    color = "Temp Deviation",
    caption = caption_w_logo(source = "Data Source")
  ) +
  guides(color = guide_colorbar(barwidth = 10, barheight = .7, reverse = T))
```

Example Using the PW Blue Palette
```{r}
# Calculate average Ozone levels by Month
avg_ozone_by_month <- aggregate(Ozone ~ Month, data = airquality, FUN = mean, na.rm = TRUE)

# Create a bar plot showing average Ozone levels by Month using the PW Blue palette
jfi_plot(avg_ozone_by_month, aes(x = factor(Month), y = Ozone, fill = factor(Month))) +
  geom_col(width = .75) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  scale_fill_jfi(type = "discrete", palette = "pw_blue") +
  labs(
    title = "Average Ozone Levels by Month",
    subtitle = "Using the PW Blue palette",
    x = "Month",
    y = "Average Ozone Level",
    fill = "Month",
    caption = caption_w_logo(source = "Data Source")
  ) +
  theme(legend.position = "none")
```

Example Using the PW Red Palette
```{r}
# Apply PW red divergent color scale using temperature deviation
jfi_plot(airquality, aes(x = Wind, y = Temp, color = TempDev)) +
  geom_point(size = 2.5) +
  scale_color_jfi(type = "divergent", palette = "pw_red", midpoint = 0) +
  labs(
    title = "Wind vs Temp with Divergent Color Scale",
    subtitle = "Divergence based on temperature deviation from the mean. Using the PW Red palette",
    color = "Temp Deviation",
    caption = caption_w_logo(source = "Data Source")
  ) +
  guides(color = guide_colorbar(barwidth = 10, barheight = .7))
```

